---
title: NYC Runs Subway Challenge 2020 - Pulling data from the Strava API
author: Wal
date: '2020-06-01'
categories:
  - fitness
tags:
  - running
  - strava
lastmod: '2020-06-02T16:15:12+01:00'
featured: no
draft: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
---

I've signed up to the [NYC Runs Subway Challenge](https://nycruns.com/race/nycruns-subway-system-challenge) which involves running a total of 245 miles / 394 kilometers in the 15 weeks between Memorial Day (25th May 2020) and Labour Day (7th September 2020).

I'm targeting an average of 4 weekly runs totaling 30km per week to give me a bit of buffer (will complete in 13 weeks) in case my schedule goes awry. 

I'm tracking my progress by pulling my running data down from the strava API and monitoring it here ....


```{r data, include=FALSE, warning=FALSE}
library(tidyverse)
library(ggTimeSeries)
library(lubridate)
library(patchwork)
library(kableExtra)

plot_data <- read_csv('data/wal_strava.csv')

```


### Tracking Progress
```{r progress_chart, echo=FALSE, fig.height=10, fig.width=8, message=FALSE, warning=FALSE}
library(ggrepel)
day_chart <- plot_data %>%  
  ggplot_calendar_heatmap('start_date',
                          'distance',
                          dayBorderSize = 0.25,
                          dayBorderColour = "grey50",
                          monthBorderSize = 1, 
                          monthBorderColour = "grey50",
                          monthBorderLineEnd = "round"
  ) +
  scale_fill_gradient(low="grey80", high="darkgreen") +
  labs(title = "Weekly runs", 
       fill = "Run Distance (Km) ",
       y = NULL, x = NULL) +
  facet_null() +
  theme_minimal() +
  theme(legend.position='bottom'
        ,plot.title = element_text(color = "grey25", size = 20, face = "bold"))

daily_runs_chart <- plot_data %>%
  ggplot(aes(start_date, distance, color = distance)) + 
  geom_col() +
  scale_color_gradient(low="grey80", high="darkgreen") +
  labs(title = "Daily distance", 
       y = "Distance (km)", 
       x = NULL) +
  guides(color = FALSE) +
  theme_minimal() +
  theme(legend.position='right' 
        ,plot.title = element_text(color = "grey25", size = 20, face = "bold"))

cumsum_distance_data <- plot_data %>%
  arrange(start_date) %>%
  mutate(total_distance = cumsum(distance)) %>%
  mutate(total_distance = ifelse(start_date > today(), NA, total_distance)) 

latest_value <- cumsum_distance_data %>%
  arrange(start_date) %>%
  filter(!is.na(total_distance)) %>%
  summarise(start_date = last(start_date), 
            total_distance = last(total_distance))

target <- plot_data %>%
  mutate(daily_target = 3.75) %>%
  mutate(total_target = cumsum(daily_target))

line_chart <- cumsum_distance_data %>%
  ggplot(aes(start_date, total_distance)) + 
  geom_line(data = target, aes(x = start_date, y = total_target), color = "grey90", size = 0.5) +
  geom_line(color = "darkgreen", size = 2) +
  geom_label(data = latest_value,
             aes(label = paste0(total_distance, 'km')),
             color = "darkgreen") +
  
  geom_hline(yintercept = 394, color = "orange", size = 2) +
  labs(title = "Progress to 394km", 
       y = "Total Distance (km)", x = NULL) +
  theme_minimal() +
  theme(legend.position='bottom'
        ,plot.title = element_text(color = "grey25", size = 20, face = "bold"))

day_chart / (daily_runs_chart + line_chart) +
plot_annotation(
  title = 'NYC Subway Challenge - 394Km in 15 Weeks',
  subtitle = "Need to average 4, 7.5km runs per week to hit the target in 13 weeks ",
  theme = theme(plot.title = element_text(size = 24))
) +
  plot_layout(ncol=1,heights=c(2.5,1.5))
  
```



## Pulling activity data from the Strava API using R


The [Strava API](https://developers.strava.com/docs/reference/) is straight forward to query once you get over the authenttication hurdles. I use R to query the API to generate the the above charts.

### OAuth Authentication

To query the API you need an OAuth token. The [Strava OAuth documentation](https://developers.strava.com/docs/authentication/) is comprehensive but in summary the steps are 

1. Create a new app and generate a client_id and secret for api access (https://developers.strava.com/docs/getting-started/#account)

2. Using the httr library, generate the oauth token

```{r oauth_token, eval = FALSE}
library(httr)
app <- oauth_app("strava", <CLIENT_ID>, <SECRET>)
endpoint <- oauth_endpoint(
  request = NULL,
  authorize = "https://www.strava.com/oauth/authorize",
  access = "https://www.strava.com/oauth/token"
)

token <- oauth2.0_token(endpoint
                        , app
                        , as_header = FALSE
                        , scope = "activity:read_all")


```

### Query the Strava athlete/activities endpoint

The Strava API is paginated so you need to write code to aggregate data from multiple returned pages.


First a small function to call the API and return a data.frame of results

```{r athlete_activities_function, eval=FALSE}
# Function to call the strava v3/athlete/activities API
# This API is paginated, so you need to pass a page_number and page_size to retrieve > 1 
# page of results (default 200 results per page)
strava_athlete_activities_api <- function(oauth_token, page_number, page_size = STRAVA_API_DEFAULT_PAGE_SIZE) {
  print(paste0("Querying Strava V1 API v3/athlete/activities [page ", page_number, "]"))
  resp <- GET(
    url = "https://www.strava.com/api/v3/athlete/activities",
    config = oauth_token,
    query = list(per_page = page_size, page = page_number))
  
  if (http_type(resp) != "application/json") {
    stop("API did not return json", call. = FALSE)
    break
  }

  # Strava API returns data in JSON format, so parse it and return as a data.frame
  jsonlite::fromJSON(content(resp, "text"), flatten = FALSE)
}
```


Then a loop to keep calling the API until no more results are available


```{r call_api_loop, eval=FALSE}

# A list to store the returned data into
data_pages <- list()

# Start with page #1
page_number <- 1

repeat {
  
# Query API
  data_pages[[page_number]] <- strava_athlete_activities_api(token, page_number)
  
  # Count number of records returned
  records_returrned <- data_pages[[page_number]] %>% nrow()
  print(paste0('Retrieved ', records_returrned, ' records'))
  
  # Continue to the next page if the number of returned records matches the page size
  if (records_returrned < STRAVA_API_DEFAULT_PAGE_SIZE) {
    print("All records returned, exiting")
    break
  } else {
    print("checking for more records ..")
    page_number <- page_number + 1
  }
}

# Combine the list of data frames into a single dataframe of all collected pages
data <- rbind_pages(data_pages)
print(paste0('returned a total of ', data %>% nrow(), ' records'))

```

You will now have the full history of your Strava activities!

Full script code is [here](https://github.com/wal/stravaR/blob/master/R/download_strava.R)


